---
title: "Congestion Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import packages

If you've never imported before, you'll need to run `install.packages("name of package here")` once.  After that, the packages will live in your local library.  All you need to do to import them is run the code chunk below.

```{r}
library(data.table)
library(tidyverse)
library(plotly)
library(extrafont)
library(sf)
library(leaflet)
extrafont::loadfonts()
```

# Import StreetLight (SL) data

These two csv files were created from two different SL projects.  We used two projects because the maximum zones (in this case, a zone = a road segment) SL permits in a single congestion analysis project is 5,000.

```{r}
congestion_1 <- fread("StL Tables/82850_CongAnal_1_4500_92519_sa_all.csv")
congestion_2 <- fread("StL Tables/82851_CongAnal_4501_8995_92519_sa_all.csv")

# Combine tables to create one table
congestion <- bind_rows(congestion_1, congestion_2)

# Make codes tables
day_types <- congestion %>%
  select(`Day Type`) %>%
  unique()

day_parts <- congestion %>%
  select(`Day Part`) %>%
  unique()

# Select relevant variables & rename to names that are more convenient to work with; create day_type_code and day_part_code variables so it's not necessary to type out entire descripter when filtering
congestion_tidy <- congestion %>%
  dplyr::select(`Zone ID`, `Day Type`, `Day Part`, `Avg All Segment Speed (mph)`) %>%
  rename(id = `Zone ID`,
         day_type = `Day Type`,
         day_part = `Day Part`,
         avg_spd = `Avg All Segment Speed (mph)`) %>%
  separate(day_type, into = c("day_type_code", "day_type_desc"), sep = 1) %>%
  separate(day_part, into = c("day_part_code", "day_part_desc"), sep = 2) %>%
  mutate(day_type_desc = str_replace(day_type_desc, pattern = ": ", replacement = ""),
         day_part_desc = str_replace(day_part_desc, pattern = ": ", replacement = ""))

# Select min speed for weekdays
congestion_wkday <- congestion_tidy %>%
  filter(day_type_code == 1 & day_part_code != "00" & day_part_code != "02" & day_part_code != "09" & day_part_code != "19") %>% # Don't include all day, off-peak, peak AM, or peak PM
  group_by(id) %>%
  mutate(min_spd = min(avg_spd)) %>%
  ungroup() %>%
  filter(avg_spd == min_spd) # Note that some segments have multiple hours with the same speed that is the min - will cause duplicates

# Look at duplicates (hours with same speed as min)
congestion_dups <- congestion_wkday %>%
  group_by(id) %>%
  count()

# Join back counts of hours to detailed dataset
congestion_check <- left_join(congestion_wkday, congestion_dups, by = "id")

# See how many times the overnight hours appear as the minimum speed (and are the only hour that occurs)
check_plot <- congestion_check %>%
  mutate(day_part_code = as.numeric(day_part_code)) %>%
  mutate(min_hour_cat = ifelse(n == 1, "One hour is min", "Multiple hours are min")) %>%
  ggplot(aes(fct_reorder(day_part_desc, day_part_code), fill = min_hour_cat)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values = c("#FF99FF", "#9999FF")) +
  labs(x = "",
       fill = "") +
  theme(text = element_text(family = "Tw Cen MT"),
        legend.position = "bottom",
        axis.ticks = element_blank(),
        panel.background = element_rect(fill = "white"),
        panel.grid = element_blank())

check_plot

ggsave("Methodology check.png", width = 10, height = 6)

ggplotly(check_plot)


```

# Take a look at where segments with 2+ hours with same min speed are located

```{r}
# Create summarized dataset that contains min speed, hours with that speed, and zone ID
congestion_wkday_id <- congestion_wkday %>%
  group_by(id, min_spd) %>%
  summarize(hour_or_hours = paste(day_part_desc, collapse = ", "))

# Add variable stating number of hours tied for min
congestion_full <- left_join(congestion_wkday_id, congestion_dups, by = "id") %>%
  rename(n_min_hours = n)

# Import shapefiles
con_shp_1 <- st_read("Shapefiles/82850_CongAnal_1_4500_92519_segment_line.shp")
con_shp_2 <- st_read("Shapefiles/82851_CongAnal_4501_8995_92519_segment_line.shp")

congestion_shp_1 <- left_join(con_shp_1, congestion_full, by = "id")
congestion_shp_2 <- left_join(con_shp_2, congestion_full, by = "id")

congestion_shp <- rbind(congestion_shp_1, congestion_shp_2) #do NOT use bind_rows

congestion_lab <- congestion_shp %>%
  mutate(popup = paste0("Number of hours tied for minimum speed: <strong>", n_min_hours, "</strong><br> Min speed: <strong>", min_spd, "</strong"))

labs <- as.list(congestion_lab$popup)

congestion_shp %>%
  leaflet() %>%
  addPolygons(color = ~colorNumeric("PuBu", n_min_hours)(n_min_hours),
      label = lapply(labs, htmltools::HTML),
      labelOptions = labelOptions(style = list("font-family" = "Segoe UI Light",
                                               "font-size" = "14px"))) %>%
  addProviderTiles(providers$CartoDB.Positron)

#st_write(congestion_shp, "Congestion Method Check.shp", "ESRI Shapefile Driver")
```

# Take a look at segments where congestion is in the off-peak

```{r}
con_wkday_1 <- left_join(con_shp_1, congestion_wkday, by = "id")
con_wkday_2 <- left_join(con_shp_2, congestion_wkday, by = "id")

congestion_wkday_shp <- rbind(con_wkday_1, con_wkday_2) #do NOT use bind_rows

st_write(congestion_wkday_shp, "Congestion Method Check on Off-peak.shp", "ESRI Shapefile Driver")

```

# Congestion (without hour designation) + free-flow speed

```{r}
# Note only 8,910 segments have weekday speeds for some hour of the day
congestion_wkday_unique <- congestion_wkday %>%
  dplyr::select(id, min_spd) %>%
  unique()

all_ids <- as_tibble(c(1:8995)) %>%
  rename(id = value)

missing_con_spds <- anti_join(all_ids, congestion_wkday_unique, by = "id")

# Note only 8,734 segments have speeds recorded for off-peak
flow_wkday <- congestion_tidy %>%
  filter(day_type_code == "1" & day_part_code == "02") %>%
  dplyr::select(id, avg_spd) %>%
  rename(off_pk_spd = avg_spd) %>%
  unique()

missing_flow_spds <- anti_join(all_ids, flow_wkday, by = "id")

flow_con <- full_join(congestion_wkday_unique, flow_wkday, by = "id")

ref_spd <- flow_con %>%
  mutate(perc_congested = min_spd/off_pk_spd*100)

ref_wkday_1 <- left_join(con_shp_1, ref_spd, by = "id")
ref_wkday_2 <- left_join(con_shp_2, ref_spd, by = "id")

congestion_flow_shp <- rbind(ref_wkday_1, ref_wkday_2)
st_write(congestion_flow_shp, "Shapefiles/Congestion_Flow_All_Hours.shp", "ESRI Shapefile Driver")
write_csv(missing_con_spds, "Segment IDs Missing Weekday Speed.csv")
write_csv(missing_flow_spds, "Segment IDs Missing Off-Peak Weekday Speeds.csv")

```


